### 2022-09-19 draft.tex
Deploying to gh-pages from @ PoisotLab/ms_interpreting_probabilities@80025a5d21ec265f30875f53df2db7aef694f593 🚀

\hypertarget{introbackground}{%
\section{Intro/Background}\label{introbackground}}

Why it is useful think about interactions as probabilistic event An
interaction is probabilistic since two species `meeting' does not mean
that an interaction will occur e.g.~a lion crossing paths with a gazelle
does not mean predation will happen but is contextual on the
physiological state of both the lion and the gazelle. Also, two species
co-occurring does not mean there's gonna meet (think of species relative
abundances)

Aim: Although it makes sense to think about interactions as
probabilities it is not without challenges. This paper aims to outline
some of these challenges/limitations of interpreting these probabilities

\begin{quote}
probably a dope conceptual figure {[}`scale' up a the nodes from and
individual to population to taxo group how would how we interpret these
probabilities change{]}
\end{quote}

\hypertarget{overview-of-probabilities}{%
\section{Overview of Probabilities}\label{overview-of-probabilities}}

How are we defining (in the literature) what the probability of
interaction is (there are many ways to slice this cake)? Weighted
Networks??? It might not be as intuitive as you would think/assume

\hypertarget{probabilistic-metawebs}{%
\section{Probabilistic Metawebs}\label{probabilistic-metawebs}}

What does a probability in the context of a metaweb mean? Can we turn
this into a local network realisation that is also probabilistic and
intuitive? Bayesian vs frequentist

\hypertarget{ecological-context-of-probabilisitic-interactions}{%
\section{Ecological Context of Probabilisitic
Interactions}\label{ecological-context-of-probabilisitic-interactions}}

A cautionary tale of how we define probabilities? Environmental context,
local abundance context Talk about individual scale and the population
scale (probability at the individual level vs the species level)
Taxonomic scale {[}`scale' up the nodes from an individual to population
to taxo group how would we interpret these probabilities change. How
does the aggregation change the interpretation? Does it?{]} How is it
analogous to spatial and temporal scaling (basically, all kinds of
scaling are just different ways to aggregate individuals/nodes).

@@

Note scaling can refer to both space and time Regional can inform local
but can local inform regional? Network area relationships (Ontario
lakes?? Or Alaska) Why probabilistic realised networks scale with area
but not probabilistic metawebs

\begin{quote}
empirical example figure
\end{quote}

\hypertarget{concluding-notes}{%
\section{Concluding Notes}\label{concluding-notes}}

\emph{Non-ecological Networks:} What can we learn from other
systems/fields e.g. social networks, probabilistic graph theory?

What even are the probabilities? What is the probability that we will
ever know the answer to that?

Be careful how we define probabilities. Be sure to be explicit about
these things/think about it carefully. Also, different interpretations
imply different scaling, and different ways to manipulate these numbers.
\emph{Maybe mention/thinking about workflow from metaweb to realisation}

Scores vs probabilities

\hypertarget{references}{%
\section{References}\label{references}}

>>>>

\hypertarget{introduction}{%
\section{Introduction}\label{introduction}}

\hypertarget{general-background}{%
\subsection{General background}\label{general-background}}

Why is it useful to think about interactions as probabilistic events?

\begin{itemize}
\tightlist
\item
  A biological interaction is probabilistic since two taxa co-occurring
  does not mean they are going to meet (e.g., think of species relative
  abundances). Also, two individuals meeting does not mean that an
  interaction will occur (e.g., a lion crossing paths with a gazelle
  does not mean predation).
\item
  Interaction is contextual on the environment and on the physiological
  state of both species (or individuals).
\item
  Representing ecological interactions as probabilistic events helps us
  predict networks across time and space and assess the spatial
  variability of biological interactions.
\end{itemize}

\textbf{Papers:} Poisot \emph{et al.} (2016)

\hypertarget{problem-and-objectives}{%
\subsection{Problem and objectives}\label{problem-and-objectives}}

Why should we use probabilities with caution when working with
ecological networks? What is the objective of this paper?

\begin{itemize}
\tightlist
\item
  There are different ways to define and interpret interaction
  probabilities depending on the study system and on the method used to
  build probabilistic networks.
\item
  It is important to document and define what we mean by an interaction
  probability because different definitions can have different
  ecological and statistical implications.
\item
  This paper aims to outline some of the challenges in interpreting
  interaction probabilities and propose an approach to thinking about
  them.
\end{itemize}

\hypertarget{definitions-and-interpretations}{%
\section{Definitions and
interpretations}\label{definitions-and-interpretations}}

\hypertarget{overview-of-interaction-probabilities}{%
\subsection{Overview of interaction
probabilities}\label{overview-of-interaction-probabilities}}

How are interaction probabilities defined in the literature? It might
not be as intuitive as one would think.

\begin{itemize}
\tightlist
\item
  There is a big difference in how we interpret the probability that two
  species \emph{can} interact (metaweb) and the probability that they
  \emph{will} interact (realized networks).
\item
  Interaction probabilities can be used to describe Boolean interactions
  (e.g., the probability that two species interact) and weighted
  interactions (e.g., the probability distribution of the amount of
  energy flow between two species).
\item
  In many studies, it is not obvious if their numbers represent
  interaction scores or probabilities (in the later case, it is rarely
  specified what these probabilities truly represent).
\end{itemize}

\hypertarget{probabilistic-metawebs}{%
\subsection{Probabilistic metawebs}\label{probabilistic-metawebs}}

What does a probability in the context of a metaweb mean?

\begin{itemize}
\tightlist
\item
  It means the probability that two taxa can interact, regardless of
  biological plasticity and environmental variability.
\item
  One observation is enough to set this probability to one.
\item
  Can we turn this into a local network realisation that is also
  probabilistic and intuitive?
\end{itemize}

\textbf{Papers:} Strydom \emph{et al.} (2022)

\hypertarget{probabilistic-local-networks}{%
\subsection{Probabilistic local
networks}\label{probabilistic-local-networks}}

What does a probability in the context of a local network mean? A
cautionary tale of how we define probabilities.

\begin{itemize}
\tightlist
\item
  It means the probability that two taxa will interact, regardless of
  biological plasticity and environmental variability.
\item
  What do we mean by two taxa will interact? That at least one
  individual from one group will interact with (e.g., predate) at least
  one other individual from the other group.
\item
  The probability is conditional on the environmental and local
  abundance contexts.
\item
  We should expect a certain number of interactions to occur depending
  on the probability value. This number depends on the number of trials,
  which is defined according to the ecological context (e.g.,
  environmental conditions, scale) in which probabilities were
  estimated. This is in contrast with probabilities in metawebs.
\end{itemize}

@@

\hypertarget{spatial-and-temporal-scales}{%
\subsection{Spatial and temporal
scales}\label{spatial-and-temporal-scales}}

How do interaction probabilities scale spatially and temporally?

\begin{itemize}
\tightlist
\item
  Why do probabilistic local networks scale with area and time but not
  probabilistic metawebs?
\item
  Network area relationships (e.g., Ontario lakes and Alaska datasets)
\item
  In metawebs, interaction probabilities do not scale with space and
  time.
\item
  In local networks, interaction probabilities scale with space and
  time. There are more opportunities of interactions (e.g., more
  environmental conditions) in a larger area and longer time period.
\item
  Regional networks can inform local networks but can local networks
  inform regional networks?
\end{itemize}

\textbf{Figure:} Empirical example of the association between the number
of interactions in realized local networks and the number of
interactions in the corresponding species subnetworks of regional
networks. We should expect the interaction to be linear below the 1:1
line.

\hypertarget{taxonomic-scale}{%
\subsection{Taxonomic scale}\label{taxonomic-scale}}

How do interaction probabilities scale taxonomically?

\begin{itemize}
\tightlist
\item
  Difference of interpretation between probabilities in networks at the
  individual level and at any higher taxonomic level.
\item
  How does the scale up of the nodes from an individual to population to
  taxonomic group group change our interpretation of interaction
  probabilities? How does the aggregation change the interpretation?
\item
  How is it analogous to spatial and temporal scaling? Basically, all
  kinds of scaling are just different ways to aggregate individuals or
  nodes.
\end{itemize}

\textbf{Figure:} Conceptual figure of how a scale up of the nodes from
an individual to a population to any taxonomic group change our
interpretation of the probability of interaction.

\hypertarget{concluding-remarks}{%
\section{Concluding remarks}\label{concluding-remarks}}

Here we present some advice moving forward.

\begin{itemize}
\tightlist
\item
  Non-ecological Networks: What can we learn from other systems/fields
  (e.g., social networks, probabilistic graph theory)?
\item
  What even are the probabilities? What is the probability that we will
  ever know the answer to that?
\item
  Be careful how we define probabilities. Be sure to be explicit about
  these things. Be sure to specify the type of interaction, the spatial,
  temporal, and taxonomic scale when presenting new data on interaction
  probabilities. We need better metadata documentation.
\item
  Be careful to use and manipulate interaction probabilities properly
  depending on how they were defined. Different interpretations imply
  different scaling, and thus different ways to manipulate these
  numbers.
\item
  Maybe mention thinking about a workflow to predict probabilistic local
  networks from probabilistic metawebs.
\end{itemize}

\hypertarget{references}{%
\section*{References}\label{references}}
\addcontentsline{toc}{section}{References}

\hypertarget{refs}{}
\begin{CSLReferences}{1}{0}
\leavevmode\hypertarget{ref-Poisot2016StrProa}{}%
Poisot, T., Cirtwill, A.R., Cazelles, K., Gravel, D., Fortin, M.-J. \&
Stouffer, D.B. (2016). The structure of probabilistic networks.
\emph{Methods in Ecology and Evolution}, 7, 303--312.

\leavevmode\hypertarget{ref-Strydom2022FooWeb}{}%
Strydom, T., Bouskila, S., Banville, F., Barros, C., Caron, D., Farrell,
M.J., \emph{et al.} (2022). Food web reconstruction through phylogenetic
transfer of low-rank network representation. \emph{Methods in Ecology
and Evolution}, n/a.

\end{CSLReferences}
